Precise Data Flow Analysis\\in the Presence of Correlated Method Calls

### 2014-12-11 content/appendix.tex
replacing R with R_IDE; introducing CFG abbreviation

\begin{proof}[\textbf{Proof of Lemma~\ref{lem:subsetifds}}]
  The transformation $\backCC$ is the same as $\backEq$, except that it can remove data-flow facts from the result:
  \begin{align*}
    \backCC\left(\result(\transCC_R(P))\right)(n)&=\{(n',\,D_n'^\Subset(\result(\transCC_R(P))))\ |\ n\in N^*\}(n)\\
      &=D_n^\Subset(\result(\transCC_R(P)))\\
      &\subseteq\mvp F(n)\\
      &=\result_\text{IFDS}(P)(n)\,.\qedhere
  \end{align*}
\end{proof}

@@

\begin{proof}[\textbf{Proof of Lemma~\ref{lem:sound}}] Let $\rho=\result(\transCC_R(P))$. Then
  \begin{align*}
    \backCC(\rho)(n)
    &=D_n^\Subset(\rho)\\
    &=\left\{d'\ |\ 
        d'\in\mvp F(n)\,\wedge\,\forall r\in R:\ \rho(n,\ d')(r)\ne\top_T\right\}.
  \end{align*}
  Since $\mvp F(n)=\bigsqcap_{q\in\ivp(n)}M_F(q)(\varnothing)$, and $p\in\ivp(n)$, it follows that
    \begin{align*}
      d&\in M_F(p)(\varnothing)\\&\subseteq\mvp F(n)\,.
    \end{align*}
  At the same time, for all receivers $r\in R$,
  \begin{align*}
    \rho(n,\,d)(r)
    &=\left(\bigsqcap_{q\in\ivp(n)}\mpddef q d\right)(r)\\
    &=\bigsqcap_{q\in\ivp(n)}\mpddef q d(r)\,.
  \end{align*}
  According to~Lemma~\ref{lem:sound3}, $\mpd(r)\ne\top_T$.
  Since $p\in\ivp(n)$,
  \begin{align*}
    \mpd(r)\subseteq\bigsqcap_{q\in\ivp(n)}\mpddef qd(r)\,.
  \end{align*}
  From $\bigsqcap_{q\in\ivp(n)}\mpddef qd(r)=\rho(n,\,d)(r)$ it follows that
  $\mpd(r)\subseteq\rho(n,\,d)(r)$.
  Therefore, $\rho(n,\,d)(r)\ne\top_T$, and $d\in D_n^\Subset(\rho)=\backCC(\rho)(n)$.
\end{proof}

@@

\begin{lemma}\label{lem:recindep} Let $P$ be an IFDS problem. Let $N^*$ be the supergraph for $P$, $D$ the set of data-flow facts, and $S\subseteq R$ a set of receivers.
  Then for any $n\in N^*$, $d\in D$, and receiver $r\in S$,
  \begin{equation}
    \result\left(\transCC_S(P)\right)(n,\,d)(r)=
    \result(\transCC_{\{r\}}(P))(n,\,d)(r)\,.
  \end{equation}
\end{lemma}
\begin{proof}
  According to~\eqref{eq:ideresult}, \eqref{eq:mvpdef}, and~\eqref{eq:envTransToEdgefn},
  \begin{align}
    \result\left(\transCC_S(P)\right)(n,\,d)(r)
    &=\mvp{\textsf{Env}}(n,\,d)(r)\notag\\
    &=\left(\bigsqcap_{q\in\ivp(n)}M_\textsf{Env}(q)(\Omega)(d)\right)(r)\notag\\
    &=\left(\bigsqcap_{q\in\ivp(n)}\bigsqcap_{q'\in\mathsf{RP}(q,\,d)}\ccedgefn S(q')(\topcc)\right)(r)\notag\\
    &=\bigcup_{q\in\ivp(n)}\bigcup_{q'\in\mathsf{RP}(q,\,d)}\ccedgefn S(q')(\topcc)(r)\,.\label{eq:resultThroughEdgefn}
  \end{align}
  Then from Lemma~\ref{lem:recindepedgefn},
  \begin{align*}
    \result\left(\transCC_S(P)\right)(n,\,d)(r)
    &=\bigcup_{q\in\ivp(n)}\bigcup_{q'\in\mathsf{RP}(q,\,d)}\ccedgefn{\{r\}}(q')(\topcc)(r)\\
    &=\result\left(\transCC_{\{r\}}(P)\right)(n,\,d)(r)\,.\qedhere
  \end{align*}
\end{proof}

@@

\begin{lemma}\label{lem:ccrectop}
   For an IFDS problem $P$, let $n\in N^*$ be a node, and $d\in D$ a dataflow fact such there exists a realizable path $p\in\textsf{RP}(n,\,d)$.
   Then, if there exists a receiver $r\in R$, such that
  \[
    \result\left(\transCC_{\{r\}}(P)\right)(n,\,d)(r)=\top_T\,,
  \]
  then $r\in \rcc$.
\end{lemma}
\begin{proof}
  As shown in~\eqref{eq:resultThroughEdgefn},
  \[
    \result\left(\transCC_{\{r\}}(P)\right)(n,\,d)(r)=\bigcup_{q\in\ivp(n)}\bigcup_{q'\in\mathsf{RP}(q,\,d)}\ccedgefn{\{r\}}(q')(\topcc)(r)\,.
  \]
  Since the latter is equal to $\top_T$, it follows that for each realizable path $p'$ to node $n$, $\ccedgefn{\{r\}}(p')(\top)(r)=\top_T$. According to~Lemma~\ref{lem:ccrectop}, this is only possible if $r\in\rcc$.
\end{proof}

@@

\begin{proof}[\textbf{Proof of Lemma~\ref{lem:onlycorrrec}}]
  From~\eqref{eq:ucc} we know that
  \begin{align*}
    \backCC(\result\left(\transCC_R(P)\right))
    &=\left\{(n,\,D_n^\Subset(\result(\transCC_R(P))))\ |\ n\in N^*\right\}.
  \end{align*}
  According to~\eqref{eq:dnq} and~Lemma~\ref{lem:recindep}, for a given $n\in N^*$,
  \begin{align*}
    D_n^\Subset&(\result(\transCC_R(P))))\\
    &=\left\{d\ |\ d\in\mvp F(n)\,\wedge\,\forall r\in R:\ \left\{(r,\,\result(\transCC_{\{r\}}(P))(n,\,d)(r))\ |\ r\in R\right\}(r)\ne\top_T\right\}\\
    &=\left\{d\ |\ d\in\mvp F(n)\,\wedge\,\forall\bm{r\in R}:\ \result(\transCC_{\{r\}}(P))(n,\,d)(r)\ne\top_T\right\}.
  \end{align*}

Since, according to Lemma~\ref{lem:ccrectop}, $\result(\transCC_{\{r\}}(P))(n,\,d)(r)$ can only be equal to $\top_T$ when $r\in\rcc$, we can conclude that
  \begin{align*}
    D_n^\Subset&(\result(\transCC_R(P))))\\
    &=\left\{d\ |\ d\in\mvp F(n)\,\wedge\,\forall\bm{r\in\rcc}:\ \result(\transCC_{\{r\}}(P))(n,\,d)(r)\ne\top_T\right\}\\
    &=D_n^\Subset(\result(\transCC_\rcc(P)))).
  \end{align*}

  Therefore,
  \begin{align*}
    \backCC(\result\left(\transCC_R(P)\right))
    &=\left\{(n,\,D_n^\Subset(\result(\transCC_\rcc(P))))\ |\ n\in N^*\right\}\\
    &=\backCC(\result\left(\transCC_\rcc(P)\right))\,.\qedhere
  \end{align*}
\end{proof}

>>>>

\begin{proof}[\textbf{Proof of Lemma~\ref{lem:subsetifds}}]
  The transformation $\backCC$ is the same as $\backEq$, except that it can remove data-flow facts from the result:
  \begin{align*}
    \backCC\left(\resultIDE(\transCC_R(P))\right)(n)&=\{(n',\,D_n'^\Subset(\resultIDE(\transCC_R(P))))\ |\ n\in N^*\}(n)\\
      &=D_n^\Subset(\resultIDE(\transCC_R(P)))\\
      &\subseteq\mvp F(n)\\
      &=\resultIFDS(P)(n)\,.\qedhere
  \end{align*}
\end{proof}

@@

\begin{proof}[\textbf{Proof of Lemma~\ref{lem:sound}}] Let $\rho=\resultIDE(\transCC_R(P))$. Then
  \begin{align*}
    \backCC(\rho)(n)
    &=D_n^\Subset(\rho)\\
    &=\left\{d'\ |\ 
        d'\in\mvp F(n)\,\wedge\,\forall r\in R:\ \rho(n,\ d')(r)\ne\top_T\right\}.
  \end{align*}
  Since $\mvp F(n)=\bigsqcap_{q\in\ivp(n)}M_F(q)(\varnothing)$, and $p\in\ivp(n)$, it follows that
    \begin{align*}
      d&\in M_F(p)(\varnothing)\\&\subseteq\mvp F(n)\,.
    \end{align*}
  At the same time, for all receivers $r\in R$,
  \begin{align*}
    \rho(n,\,d)(r)
    &=\left(\bigsqcap_{q\in\ivp(n)}\mpddef q d\right)(r)\\
    &=\bigsqcap_{q\in\ivp(n)}\mpddef q d(r)\,.
  \end{align*}
  According to~Lemma~\ref{lem:sound3}, $\mpd(r)\ne\top_T$.
  Since $p\in\ivp(n)$,
  \begin{align*}
    \mpd(r)\subseteq\bigsqcap_{q\in\ivp(n)}\mpddef qd(r)\,.
  \end{align*}
  From $\bigsqcap_{q\in\ivp(n)}\mpddef qd(r)=\rho(n,\,d)(r)$ it follows that
  $\mpd(r)\subseteq\rho(n,\,d)(r)$.
  Therefore, $\rho(n,\,d)(r)\ne\top_T$, and $d\in D_n^\Subset(\rho)=\backCC(\rho)(n)$.
\end{proof}

@@

\begin{lemma}\label{lem:recindep} Let $P$ be an IFDS problem. Let $N^*$ be the supergraph for $P$, $D$ the set of data-flow facts, and $S\subseteq R$ a set of receivers.
  Then for any $n\in N^*$, $d\in D$, and receiver $r\in S$,
  \begin{equation}
    \resultIDE\left(\transCC_S(P)\right)(n,\,d)(r)=
    \resultIDE(\transCC_{\{r\}}(P))(n,\,d)(r)\,.
  \end{equation}
\end{lemma}
\begin{proof}
  According to~\eqref{eq:ideresult}, \eqref{eq:mvpdef}, and~\eqref{eq:envTransToEdgefn},
  \begin{align}
    \resultIDE\left(\transCC_S(P)\right)(n,\,d)(r)
    &=\mvp{\textsf{Env}}(n,\,d)(r)\notag\\
    &=\left(\bigsqcap_{q\in\ivp(n)}M_\textsf{Env}(q)(\Omega)(d)\right)(r)\notag\\
    &=\left(\bigsqcap_{q\in\ivp(n)}\bigsqcap_{q'\in\mathsf{RP}(q,\,d)}\ccedgefn S(q')(\topcc)\right)(r)\notag\\
    &=\bigcup_{q\in\ivp(n)}\bigcup_{q'\in\mathsf{RP}(q,\,d)}\ccedgefn S(q')(\topcc)(r)\,.\label{eq:resultThroughEdgefn}
  \end{align}
  Then from Lemma~\ref{lem:recindepedgefn},
  \begin{align*}
    \resultIDE\left(\transCC_S(P)\right)(n,\,d)(r)
    &=\bigcup_{q\in\ivp(n)}\bigcup_{q'\in\mathsf{RP}(q,\,d)}\ccedgefn{\{r\}}(q')(\topcc)(r)\\
    &=\resultIDE\left(\transCC_{\{r\}}(P)\right)(n,\,d)(r)\,.\qedhere
  \end{align*}
\end{proof}

@@

\begin{lemma}\label{lem:ccrectop}
   For an IFDS problem $P$, let $n\in N^*$ be a node, and $d\in D$ a dataflow fact such there exists a realizable path $p\in\textsf{RP}(n,\,d)$.
   Then, if there exists a receiver $r\in R$, such that
  \[
    \resultIDE\left(\transCC_{\{r\}}(P)\right)(n,\,d)(r)=\top_T\,,
  \]
  then $r\in \rcc$.
\end{lemma}
\begin{proof}
  As shown in~\eqref{eq:resultThroughEdgefn},
  \[
    \resultIDE\left(\transCC_{\{r\}}(P)\right)(n,\,d)(r)=\bigcup_{q\in\ivp(n)}\bigcup_{q'\in\mathsf{RP}(q,\,d)}\ccedgefn{\{r\}}(q')(\topcc)(r)\,.
  \]
  Since the latter is equal to $\top_T$, it follows that for each realizable path $p'$ to node $n$, $\ccedgefn{\{r\}}(p')(\top)(r)=\top_T$. According to~Lemma~\ref{lem:ccrectop}, this is only possible if $r\in\rcc$.
\end{proof}

@@

\begin{proof}[\textbf{Proof of Lemma~\ref{lem:onlycorrrec}}]
  From~\eqref{eq:ucc} we know that
  \begin{align*}
    \backCC(\resultIDE\left(\transCC_R(P)\right))
    &=\left\{(n,\,D_n^\Subset(\resultIDE(\transCC_R(P))))\ |\ n\in N^*\right\}.
  \end{align*}
  According to~\eqref{eq:dnq} and~Lemma~\ref{lem:recindep}, for a given $n\in N^*$,
  \begin{align*}
    D_n^\Subset&(\resultIDE(\transCC_R(P))))\\
    &=\left\{d\ |\ d\in\mvp F(n)\,\wedge\,\forall r\in R:\ \left\{(r,\,\resultIDE(\transCC_{\{r\}}(P))(n,\,d)(r))\ |\ r\in R\right\}(r)\ne\top_T\right\}\\
    &=\left\{d\ |\ d\in\mvp F(n)\,\wedge\,\forall\bm{r\in R}:\ \resultIDE(\transCC_{\{r\}}(P))(n,\,d)(r)\ne\top_T\right\}.
  \end{align*}

Since, according to Lemma~\ref{lem:ccrectop}, $\resultIDE(\transCC_{\{r\}}(P))(n,\,d)(r)$ can only be equal to $\top_T$ when $r\in\rcc$, we can conclude that
  \begin{align*}
    D_n^\Subset&(\resultIDE(\transCC_R(P))))\\
    &=\left\{d\ |\ d\in\mvp F(n)\,\wedge\,\forall\bm{r\in\rcc}:\ \resultIDE(\transCC_{\{r\}}(P))(n,\,d)(r)\ne\top_T\right\}\\
    &=D_n^\Subset(\resultIDE(\transCC_\rcc(P)))).
  \end{align*}

  Therefore,
  \begin{align*}
    \backCC(\resultIDE\left(\transCC_R(P)\right))
    &=\left\{(n,\,D_n^\Subset(\resultIDE(\transCC_\rcc(P))))\ |\ n\in N^*\right\}\\
    &=\backCC(\resultIDE\left(\transCC_\rcc(P)\right))\,.\qedhere
  \end{align*}
\end{proof}

### 2014-12-11 content/background.tex
replacing R with R_IDE; introducing CFG abbreviation

\subsection{Terminology and Notation}
%We will start by introducing several concepts used by the IFDS and IDE analyses.
The \textit{control-flow graph} of a procedure is a directed graph whose nodes are instructions, and which contains an edge from $n_1$ to $n_2$ whenever
$n_2$ may execute immediately after $n_1$. A control-flow graph has a distinguished \textit{start node} \textsf{start}$_p$ and \textit{end node} \textsf{end}$_p$.
%and edges represent transfer of control between the instructions during an execution of the program. 
%A control-flow graph has a unique start node, $\startmain$, which is the node corresponding to the program entrypoint.
%An \textit{intra-procedural} path is a path in a control-flow graph whose nodes are in the same procedure. By contrast, an \textit{inter-procedural path} is one that contains nodes from different procedures. 
The \textit{control-flow supergraph} of a program contains the control-flow graphs of all of the procedures as
subgraphs. In addition, for each call instruction $c$, the supergraph contains a \textit{call-to-start} edge to the start node of every procedure that
may be called from $c$, and an \textit{end-to-return} edge from the end node of the procedure back to the call instruction.
\otodo{do we really need to extra "return nodes" that the IFDS/IDE papers add?}
%is a control-flow graph
%in which each procedure $p$ is augmented with an additional \textit{start node} \textsf{start}$_p$ and \textit{end node} \textsf{end}$_p$, and for each call $c_q$ to a procedure $q$, there is a \textit{call node} \textsf{call}$_{c_q}$ and subsequent \textit{return node} \textsf{return}$_{c_q}$.

@@

\commentout{
Note that $R_f$ constructs pairs of dataflow facts so that
\begin{itemize}
  \item there is always an edge $(\mathbf0,\,\mathbf0)$ corresponding to the control-flow-graph edge;
  \item if there is an edge $(\mathbf0,\,d_j)$, then there is no other edge leading to $d_j$; in particular, there is never an edge $(d_i,\,\mathbf0)$ where $d_i\ne\mathbf0$.
\end{itemize}
}

@@

\commentout{
\subsubsection{Environment Transformers}
For each node in the control-flow graph, the result of an IDE analysis computes an environment $\textsf{Env}(D,\,L)$, which is a map from data-flow facts to lattice elements.

@@

\commentout{
Given a labeled exploded supergraph, the IDE algorithm computes the environments for all nodes in the control-flow graph.

>>>>

\subsection{Terminology and Notation}
%We will start by introducing several concepts used by the IFDS and IDE analyses.
\begin{mdelete}\mtodo{We already say (but less precisely) what a CFG is in the introduction.}
The \textit{control-flow graph} of a procedure is a directed graph whose nodes are instructions, and which contains an edge from $n_1$ to $n_2$ whenever
$n_2$ may execute immediately after $n_1$. A CFG has a distinguished \textit{start node} \textsf{start}$_p$ and \textit{end node} \textsf{end}$_p$.
\end{mdelete}
%and edges represent transfer of control between the instructions during an execution of the program. 
%A control-flow graph has a unique start node, $\startmain$, which is the node corresponding to the program entrypoint.
%An \textit{intra-procedural} path is a path in a control-flow graph whose nodes are in the same procedure. By contrast, an \textit{inter-procedural path} is one that contains nodes from different procedures. 
The \textit{control-flow supergraph} of a program contains the CFGs of all of the procedures as
subgraphs. In addition, for each call instruction $c$, the supergraph contains a \textit{call-to-start} edge to the start node of every procedure that
may be called from $c$, and an \textit{end-to-return} edge from the end node of the procedure back to the call instruction.
\otodo{do we really need to extra "return nodes" that the IFDS/IDE papers add?}
%is a control-flow graph
%in which each procedure $p$ is augmented with an additional \textit{start node} \textsf{start}$_p$ and \textit{end node} \textsf{end}$_p$, and for each call $c_q$ to a procedure $q$, there is a \textit{call node} \textsf{call}$_{c_q}$ and subsequent \textit{return node} \textsf{return}$_{c_q}$.

@@

\commentout{
Note that $R_f$ constructs pairs of dataflow facts so that
\begin{itemize}
  \item there is always an edge $(\mathbf0,\,\mathbf0)$ corresponding to the CFG edge;
  \item if there is an edge $(\mathbf0,\,d_j)$, then there is no other edge leading to $d_j$; in particular, there is never an edge $(d_i,\,\mathbf0)$ where $d_i\ne\mathbf0$.
\end{itemize}
}

@@

\commentout{
\subsubsection{Environment Transformers}
For each node in the CFG, the result of an IDE analysis computes an environment $\textsf{Env}(D,\,L)$, which is a map from data-flow facts to lattice elements.

@@

\commentout{
Given a labeled exploded supergraph, the IDE algorithm computes the environments for all nodes in the CFG.
